DRAFT: Week 3: Degree and Centrality

### 2022-03-08 docs/3. SNA - Assignments/DRAFT_Week3-Distill-Blog_Degree-and-Centrality.Rmd
3.6.22

## To start:

### I'm curious to see what sort of relationships exist between the components of [Smuggler's Cove](https://www.smugglerscovesf.com/#about-section) Tiki Cocktail Recipies?

@@

    vertices = 274

    -   Nodes:

        Drink Name

        Drink Ingredient

@@

@@

```{=html}
<!-- -->
```
-   Reach-ability:

@@

-   Network Density

-   Degree: Network vs Node

-   Degree Centrality: Network vs Node

>>>>

### Which drink ingredients in [Smuggler's Cove](https://www.smugglerscovesf.com/#about-section)'s have the most connections and are the most central in this network of Tiki Cocktail Recipes?

@@

    -   Nodes = 274

        -   Drink Name

        -   Drink Ingredient

@@

    Edge attributes: Ingredient Type

    > 10 most common values are:
    >
    > JUICE : 135 , RUM: 125, SYRUP: 115, LIQUEUR
    > : 84 , BITTERS: 54 ,
    >
    > SODA : 28 , SPICE: 19, GIN: 12, BRANDY: 9,
    > FRUIT: 7

@@

    Network Transitivity is either 1 or 0. Statnet
    read in the edgelist as "directed", which
    igraph is "undirected" .

    According to statnet the network is entirely
    transitive because the graph is composed
    entirely of dyads and potential traids

    Igraph: Transitivity is near 0 (0.018) because
    network is composed of 627 mutual ties, no
    Asymetic ties, and 36,729 null ties.

    statnet: 0 mutual, 671 asymetric, 36,730
    potential ties

-   Reach-ability:

@@

1 component in graph

all components are connected

### Network Density: Low Transitivity == Low Density

*The density of a network is defined as the
proportion of ties present out of all possible
ties in a network.*

statnet:

-   Network Density: 0

-   gDen() - Edgelist Network Density: 0.0089

igraph:

-   Graph Density: 0.0179

-   Edge Density: 0.0179

### Degree: Network vs Vertex

\*\*Vertex Degree\*\* measures an individual
node's centrality o; each vertex in a network may
have a different node degree.

#### Degree: Network vs Node

[Network Degree Summary Stats]{.ul}

+-----+---+------+---+------+---+
| deg |   | inde |   | o    |   |
| ree |   | gree |   | utde |   |
|     |   |      |   | gree |   |
+=====+===+======+===+======+===+
| M   | 0 | Min. | 0 | Min. | 0 |
| in. |   |      | . |      |   |
|     |   |      | 0 |      |   |
|     |   |      | 0 |      |   |
|     |   |      | 0 |      |   |
+-----+---+------+---+------+---+
| 1st | 0 | 1st  | 0 | 1st  | 0 |
| Qu. |   | Qu.  |   | Qu.  |   |
+-----+---+------+---+------+---+
| Med | 1 | Me   | 1 | Me   | 0 |
| ian |   | dian |   | dian |   |
+-----+---+------+---+------+---+
| M   | 2 | Mean | 2 | Mean | 2 |
| ean | . |      | . |      | . |
|     | 4 |      | 4 |      | 4 |
|     | 5 |      | 4 |      | 4 |
|     |   |      | 9 |      | 9 |
+-----+---+------+---+------+---+
| 3rd | 2 | 3rd  | 2 | 3rd  | 5 |
| Qu. |   | Qu.  |   | Qu.  |   |
+-----+---+------+---+------+---+
| M   | 6 | Max. | 6 | Max. | 1 |
| ax. | 2 |      | 2 |      | 1 |
+-----+---+------+---+------+---+
|     |   |      |   |      |   |
+-----+---+------+---+------+---+

[Vertex Degree Summary Stats]{.ul}

+-----------+-------------+-----------------+
| degree    | indegree    | outdegree       |
+===========+=============+=================+
| Min:      | Min: 0.000  | Min: 0.000      |
| 1.000     |             |                 |
+-----------+-------------+-----------------+
| 1st Qu:   | 1st Qu:     | 1st Qu: 0.000   |
| 1.000     | 0.000       |                 |
+-----------+-------------+-----------------+
| Median:   | Median:     | Median: 0.000   |
| 4.000     | 1.000       |                 |
+-----------+-------------+-----------------+
| Mean:     | Mean: 2.449 | Mean: 2.449     |
| 4.898     |             |                 |
+-----------+-------------+-----------------+
| 3rd Qu:   | 3rd Qu:     | 3rd Qu: 5.000   |
| 6.000     | 2.000       |                 |
+-----------+-------------+-----------------+
| Max:      | Max: 62.000 | Max: 11.000     |
| 62.000    |             |                 |
+-----------+-------------+-----------------+

[Network Degree Distribution Graphs]{.ul}

#### Degree Centrality: Network vs Node

Vertex Degree Centrality:

### 2022-03-08 docs/3. SNA - Assignments/sna_Week3/DRAFT-week3-Assignment.Rmd
3.6.22 time for a walk
3.6.22

#   summarize numeric network attribute
  summary(snEL%v%"name")
  summary(snEL%v%"ingNam")
  summary(snEL%v%"dnkNam")
  summary(snEL%e%"ingTyp")

@@

summary(
  graph.density(igEL)
)

@@

summary(
   edge_density(igEL)
)

@@

 igEL_randoGraph <- erdos.renyi.game(
    n = igEL_Nod, 
    p.or.m = igEL_Den, 
    directed = T) 
      # where n is the number of nodes, 
      # p.or.m is the probability of drawing an edge, 
      #directed is whether the network is directed or not
 class(igEL_randoGraph)
 summary(igEL_randoGraph)

 print_all(igEL_randoGraph)

```{r iGraph: GRAPH: Dyad Census, echo=TRUE}

@@

```{r iGraph: Triad Census, echo=TRUE}
library(igraph)
#Classify all Triad in the network

sum(
  triad_census(igEL)
)

igEL_triList <- 
  triad_census(igEL)

@@

View(
geodist(snEL))

@@

summary(
  reachability(
    snEL,
    geodist.precomp= gd_snEL,
    return.as.edgelist=F,
    na.omit=F)
)
```

@@

gplot(kpath.census(snEL, 
             maxlen = 10, 
             mode = "graph",
             tabulate.by.vertex = T, 
             path.comembership = c("bylength"), # c("none", "sum","bylength"),
             dyadic.tabulation = c("sum") # c("none", "sum", "bylength"))
))
              ##**"sum"* returns a vertex by vertex matrix of source/destination path counts
             ##**"bylength"* disaggregates these counts by path length. 
             ##**"none"* disables this computation.

@@

view(
  distance_table(igEL, directed = F)
)

@@

@@

 igraph::components(igEL)$no # mode = c("weak", "strong"))
      ##**NOTE:** $no prevents a list of components from being printed
summary(
  igraph::components(igEL,
                     mode = c("weak"))
)

summary(
  igraph::components(igEL,
                     mode = c("strong"))
)
 # Size of each component

@@

summary(
component.largest(
 snEL, 
  connected=c("weak"),
  result = c("membership"), # "membership", "graph")
  return.as.edgelist = T)
)

#  computes the component structure of the input network, and returns a vector

summary(
component.size.byvertex(snEL)
)

summary(
component.size.byvertex(
 snEL, 
  connected = c("unilateral")) # "strong", "weak", "unilateral", "recursive"
)

@@

  view(summary(cl_snEL))
  View(cl_snEL)

gplot(cl_snEL)

@@

    network.density(
     snEL, 
     na.omit = T, # Exclude missing edges
     discount.bipartite = FALSE)

 # computes the density of  graphs and adjusts for the type of graph 
  gden(snEL)

  gden(
    snEL, 
       g=NULL, 
       diag=FALSE, # Set to TRUE if graph contains loops
       mode="graph", # digraph or graph
       ignore.eval=FALSE)

@@

## Network & Vertex Degree

### **use the network degree command to:**

-   compute network summary statistics (such as
    average network degree);
-   graph the network degree distribution; or
-   identify particular nodes (such as the five
    nodes with the most contacts). ##\# **Vertex
    degree** is a description of an individual
    node; each vertex in a network may have a
    different node degree.

###*Node Degree:* node degree is one measure of
node centrality or popularity; Node degree cis a
count of the relationships that a node is involved
in or the count of the edges that are incident to
a vertex

### Statnet: Network & Vertex Degree

#### Network Degree
```{r Statnet: Network Degree, echo=TRUE}
library(statnet)
library(sna)

 # statnet: Calculate average network degree
summary(  
  sna::degree(snEL))

summary(
  sna::degree(snEL, 
             g=1, 
             nodes=NULL, 
             gmode="digraph", #directed;
             diag=FALSE, # contains loops
             tmaxdev=FALSE, 
             cmode="indegree", 
             rescale=FALSE, 
             ignore.eval=FALSE)) # ignores loops or self-nominations 
summary(
  snEL_vd <- sna::degree(snEL, 
             g=1, 
             nodes= NULL, 
             gmode="graph", #undirected
             diag=TRUE, #includes diagonals 
             tmaxdev=F, 
             cmode="freeman", 
             rescale=F, #  scores rescaled to sum to 1.
             ignore.eval=FALSE))

@@

```{r Statnet: Vertex Degree (Directed Network), echo=TRUE}
library(statnet)
library(sna)

     ## create a dataset of vertex names and degree:
#####
snEL_nDeg <- data.frame(
  name=snEL%v%"vertex.names",
  degree=sna::degree(snEL))

@@

@@

#### Network Degree Summary Statistics

@@

#### Network PLOT: Degree Distribution

@@

```{r iGraph: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree
 # provides node name along with node degree

@@

We can break large social networks down into their
constituent parts. These constituent parts are
referred to as "motifs". The most basic motif
consists of two nodes and is called a dyad. Edges
in a network signify the presence or absence of
dyadic relations. It follows that a dyad in an
undirected network can have two unique
configurations: connected or disconnected; and
three unique configurations in a directed network
(mutual, assymetric, and null

@@

summary(
  centr_degree(igEL)$centralizatio)

>>>>

#   summarize numeric network attribute
  summary(snEL%v%"vertex.name")
  #summary(snEL%v%"ingNam")
  #summary(snEL%v%"dnkNam")
  #summary(snEL%e%"ingTyp")

@@

@@

@@

 igEL_randoGraph <- erdos.renyi.game(
    n = igEL_Nod, 
    p.or.m = igEL_Den, 
    directed = T) 
      # where n is the number of nodes, 
      # p.or.m is the probability of drawing an edge, 
      #directed is whether the network is directed or not
 class(igEL_randoGraph)

 graph.density(igEL_randoGraph)
 edge_density(igEL_randoGraph)

```
```{r iGraph: GRAPH: Dyad Census, echo=TRUE}

@@

```{r iGraph: Triad Census, echo=TRUE}
library(igraph)
#Classify all Triad in the network
igEL_triList <- triad_census(igEL)
igEL_triList

sum(
  triad_census(igEL)
)
summary(igEL_triList)

@@

summary(
geodist(snEL))

@@

view(
  reachability(
    snEL,
    geodist.precomp= gd_snEL,
    return.as.edgelist=F,
    na.omit=F)
)
```

@@

gplot(kpath.census(snEL, 
             maxlen = 5, 
             mode = "graph",
             tabulate.by.vertex = T, 
             path.comembership = c("bylength"), # c("none", "sum","bylength"),
             dyadic.tabulation = c("sum") # c("none", "sum", "bylength"))
))
              ##**"sum"* returns a vertex by vertex matrix of source/destination path counts
             ##**"bylength"* disaggregates these counts by path length. 
             ##**"none"* disables this computation.

@@

  distance_table(igEL, directed = F)

@@

# Size of each component

@@

 igraph::components(igEL)$no# mode = c("weak", "strong"))
      ##**NOTE:** $no prevents a list of components from being printed
summary(
  igraph::components(igEL,
                     mode = c("weak"))
)

summary(
  igraph::components(igEL,
                     mode = c("strong"))
)

@@

component.largest(
 snEL, 
  connected=c("weak"),
  result = c("membership"), # "membership", "graph")
  return.as.edgelist =FALSE)

#  computes the component structure of the input network, and returns a vector 
sum(
component.size.byvertex(snEL)
)

sum(
component.size.byvertex(
  snEL, 
  connected = c("unilateral")) # "strong", "weak", "unilateral", "recursive"
)

@@

# View(cl_snEL)

```
gplot(cl_snEL)

@@

    network.density(
     snEL, 
     na.omit = T, # Exclude missing edges
     discount.bipartite=F)

 # computes the density of  graphs and adjusts for the type of graph 
  gden(snEL,
       diag = F)

  gden(
    snEL, 
       g=NULL, 
       diag=T, # Set to TRUE if graph contains loops
       mode="graph", # digraph or graph
       ignore.eval=FALSE)

 gden(snEL,
      diag = T,
      mode = "digraph")

@@

## Degree: Network & Vertex
### **use the network degree command to:**

-   compute network summary statistics (such as
    average network degree);
-   graph the network degree distribution; or
-   identify particular nodes (such as the five
    nodes with the most contacts). 
    **Vertex degree** is a description of an individual
    node; each vertex in a network may have a
    different node degree.

*Node Degree:* node degree is one measure of
node centrality or popularity; Node degree cis a
count of the relationships that a node is involved
in or the count of the edges that are incident to
a vertex

### Statnet: Network & Vertex Degree
#### Statnet: Network Degree
```{r Statnet: Network Degree, echo=TRUE}
library(statnet)
library(sna)

  sna::degree(snEL)$no
summary(
  snEL_vd <- sna::degree(snEL, 
             g=1, 
             nodes= NULL, 
             gmode="graph", #undirected
             diag=TRUE, #includes diagonals 
             tmaxdev=F, 
             cmode="freeman", 
             rescale=F, #  scores rescaled to sum to 1.
             ignore.eval=FALSE))

summary(
  sna::degree(snEL, 
             g=1, 
             nodes=NULL, 
             gmode="digraph", #directed;
             diag=FALSE, # contains loops
             tmaxdev=FALSE, 
             cmode="indegree", 
             rescale=FALSE, 
             ignore.eval=FALSE)) # ignores loops or self-nominations

summary(
  sna::degree(snEL, 
             g=1, 
             nodes=NULL, 
             gmode="digraph", #directed;
             diag=FALSE, # contains loops
             tmaxdev=FALSE, 
             cmode="outdegree", 
             rescale=FALSE, 
             ignore.eval=FALSE)) # ignores loops or self-nominations

@@

#### Statnet: Vertex Degree
```{r Statnet: Vertex Degree (Directed Network), echo=TRUE}
library(statnet)
library(sna)

     ## create a dataset of vertex names and degree:
#####

@@

snEL_nDeg <- data.frame(
  name=snEL%v%"vertex.names",
  degree=sna::degree(snEL))

@@

@@

@@

```{r iGraph: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree

 # provides node name along with node degree
summary(igraph::degree(igEL))

@@

A dyad in an undirected network can have two unique configurations: connected or disconnected; 
and three unique configurations in a directed network
(mutual, assymetric, and null

@@

summary(
  centr_degree(igEL)$centralization)
